---
title: "Carry_Strategy"
author: "Maximilian Michlits"
header-includes:
   - \usepackage{bbm}
date: "r Sys.Date()"
output: html_document
---

To Do's:
Set up proper data loading
Inlcude 3 factors that Simion sent (in icloud -> Studium -> Simion, Giorgia "Questions")
Statistical interactions


Starting off by loading libraries and the data.
```{r, include = FALSE}
#load libraries
library(dplyr) #general use
library(ggplot2) #plots

#Debug T = on/ F = off
DEBUG <- FALSE
```

We want to calculate the carry of each currency. The Carry $C$, of currency $i$ against USD, at time $t$ and horizon $h\in (0,12)$ in months (the time span into the future for the forward) can be computed as follows: $C_{i_t}^h=Forward_{i_t}^h-Spot_{i_t}^h$

Luckily this is the same as forward spreads which is the standard way of downloading forward data via Bloomberg. Therefore, no further calculations are necessary to find carry.

I use monthly data for the following to reduce noise and speed up computation time. This methodology is further validated by application constraints, as reallocating more often results in extremly high trading costs.

Other factors we load are:
liquidity = (bid-ask) / ((bid+ask)/2), capturing relative spread relative to price.
VIX, which is the implied volatility of the S&P500 for one year, basically markets anticipation of risk for the next year.
```{r}
#load data sets.
load("./Data/fx_factors.RData")

fx_factors <- fx_factors %>%
    select(date, Currency, log_return, Liquidity, everything())

risk_free_rate <- load("./Data/eur_risk_free_rate.RData")

#tmp fixes
fx_factors <- fx_factors %>%
  mutate(log_return = lead(log_return, 1))

fx_factors$REER_historic_mean_ratio <- 1/fx_factors$REER_historic_mean_ratio

#fx_factors <- fx_factors %>%
#  filter(date <= "2019-12-31")

#display df
head(fx_factors, 20)
```

#Building models
##Factor Portfolio Construction
We build factor portfolios for all factors by defining a function that takes a factor and puts together the n best and worst performing positions under the given factor. The weighting scheme will be either 1/n or signal weighted.

I also want to adjust the weighting to the risk contribution of each currency. https://books.google.at/books?hl=de&lr=&id=3_tZCgAAQBAJ&oi=fnd&pg=PP1&dq=weighting+investments+based+on+risk+in+a+factor+strategy&ots=F6YX8n9y4-&sig=CVb0Q8hxCU7haLavyFrsi-Admfw#v=onepage&q=weighting%20investments%20based%20on%20risk%20in%20a%20factor%20strategy&f=false P.18. maybe? We also aim to use Markowitz for the final portfolio composition because every factor portfolio should have some expected return and some variance. This can be used for the optimization.

Signal weighting takes sum of the factor levels for each currency for the n best or worst currencies and then takes their individual factor level relative to the overall factor level.

```{r}
# Define the within-portfolio weighting schemes via their own function.
weighting <- function(data.timestamp, n = 1, mode = "1/n", strategy = "long_short", liquidity_threshold = NULL) {
  colnames(data.timestamp)[5] <- "fx_factor"
  
  # Filter out illiquid currencies if threshold is specified
  if (!is.null(liquidity_threshold)) {
    data.timestamp <- data.timestamp %>%
      filter(Liquidity <= liquidity_threshold)
  }
  
  # Generate long leg
  long_leg <- data.timestamp %>%
    arrange(desc(fx_factor)) %>%
    head(n) %>%
    mutate(
      weight = if (mode == "1/n") 1 / n else fx_factor / sum(fx_factor),
      scaled_log_return = log_return * weight,  # Scale log returns by weights
      leg = "long"
    )
  
  # Generate short leg
  short_leg <- data.timestamp %>%
    arrange(fx_factor) %>%
    head(n) %>%
    mutate(
      weight = if (mode == "1/n") -1 / n else fx_factor / sum(fx_factor),
      scaled_log_return = log_return * weight,  # Scale log returns by weights
      leg = "short"
    )
  
  # Adjust based on strategy
  if (strategy == "long_only") {
    short_leg <- data.frame()
  } else if (strategy == "short_only") {
    long_leg <- data.frame()
  }
  
  # Combine legs into one data frame
  combined_legs <- bind_rows(long_leg, short_leg) %>%
    select(Currency, date, scaled_log_return, weight, leg)
  
  colnames(combined_legs)[3] <- "log_return"
  
  return(combined_legs)
}

# Function to construct portfolio
portfolio_construction <- function(data, n = 1, mode = "1/n", strategy = "long_short", liquidity_threshold = NULL, factor_name = "") {
  # Initialize portfolio data frame
  factor_portfolio <- data.frame(
    Currency = character(),
    date = as.Date(character()),
    log_return = numeric(),
    weight = numeric(),
    leg = character(),
    factor_name = character()
  )
  
  # Loop through unique dates and construct portfolios
  for (current_date in unique(data$date)) {
    weighted_data <- weighting(
      data.timestamp = data %>% filter(date == current_date),
      n = n,
      mode = mode,
      strategy = strategy,
      liquidity_threshold = liquidity_threshold
    )
    if (nrow(weighted_data) > 0) {
      weighted_data <- weighted_data %>%
        mutate(factor_name = factor_name)
      factor_portfolio <- bind_rows(factor_portfolio, weighted_data)
    }
  }
  
  return(factor_portfolio)
}
```

We need a few helper functions.
```{r}
# Helper functions for factor log returns of a portfolio
compute_factor_log_return <- function(factor_portfolio, start_date = NULL, end_date = NULL) {
  factor_log_return <- factor_portfolio

# Filter by start_date if provided
if (!is.null(start_date)) {
  factor_log_return <- factor_log_return %>%
    filter(date >= as.Date(start_date))
}

# Filter by end_date if provided
if (!is.null(end_date)) {
  factor_log_return <- factor_log_return %>%
    filter(date <= as.Date(end_date))
}

# Group and summarize
factor_log_return <- factor_log_return %>%
  group_by(date) %>%
  summarise(dates_log_return = sum(log_return, na.rm = TRUE)) %>%
  ungroup()
  
  return(factor_log_return)
}

compute_cumulative_log_return <- function(factor_portfolio, start_date = NULL, end_date = NULL) {
  filtered_log_return <- compute_factor_log_return(factor_portfolio)
  
  # Filter by start_date if provided
  if (!is.null(start_date)) {
    filtered_log_return <- filtered_log_return %>%
      filter(date >= as.Date(start_date))
  }
  
  # Filter by end_date if provided
  if (!is.null(end_date)) {
    filtered_log_return <- filtered_log_return %>%
      filter(date <= as.Date(end_date))
  }
  
  # Compute cumulative log return
  cum_log_return <- filtered_log_return %>%
    mutate(cum_log_return = cumsum(dates_log_return)) %>%
    select(-dates_log_return) # Optionally remove dates_log_return
  
  return(cum_log_return)
}

# Helper function for plotting cumulative log returns
plot_factor_portfolio <- function(factor_portfolio, factor_name, line_color = "blue", y_min = NULL, y_max = NULL) {
  cum_log_return <- compute_cumulative_log_return(factor_portfolio)
  
  # Create the base plot
  plot <- ggplot(cum_log_return, aes(x = date, y = cum_log_return)) +
    geom_line(color = line_color, size = 1) +
    labs(
      title = paste("Cumulative Log Returns for", factor_name, "Over Time"),
      x = "Date",
      y = "Cumulative Log Return"
    ) +
    theme_minimal() +
    theme(
      plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
      axis.title = element_text(size = 12),
      axis.text = element_text(size = 10)
    )
  
  # Add y-axis limits if specified
  if (!is.null(y_min) & !is.null(y_max)) {
    plot <- plot + ylim(y_min, y_max)
  }
  
  return(plot)
}

# Helper function for portfolio risk measures
risk_measure_summary <- function(
    factor_portfolio,
    risk_free_rate,
    start_date = NULL,
    end_date = NULL){
  #pull underlying data
  simple_log_return <- compute_factor_log_return(factor_portfolio, start_date, end_date)
  cum_log_return <- compute_cumulative_log_return(factor_portfolio, start_date, end_date)
  num_periods <- nrow(simple_log_return)
  risk_free_rate <- (1 + mean(risk_free_rate$Level, na.rm = TRUE))^(1 / 12) - 1
  
  
  #return measures
  arithmetic_mean_return <- sum(simple_log_return$dates_log_return, na.rm = TRUE) / num_periods
  geom_mean_return <- exp(arithmetic_mean_return) - 1
  
  #vola measures
  volatility <- sd(simple_log_return$dates_log_return, na.rm = TRUE)
  
  #SR measure
  sharpe_ratio <- (arithmetic_mean_return - risk_free_rate)/volatility
  
  return(data.frame(
    arithmetic_mean_return = arithmetic_mean_return,
    geom_mean_return = geom_mean_return,
    volatility = volatility,
    sharpe_ratio = sharpe_ratio
  ))
}

#generate multi factor combinations
generate_combinations <- function(factor_list, min_constituents = 1, max_constituents = NULL) {
  # Set max_constituents to the length of the factor list if not provided
  if (is.null(max_constituents)) {
    max_constituents <- length(factor_list)
  }
  
  # Validate input
  if (min_constituents > max_constituents) {
    stop("min_constituents cannot be greater than max_constituents")
  }
  
  # Generate combinations for the specified range
  combinations_list <- unlist(
    lapply(min_constituents:max_constituents, function(k) combn(factor_list, k, simplify = FALSE)),
    recursive = FALSE
  )
  
  return(combinations_list)
}
```

With this loop we can now construct each factor portfolio at once. There is a bug that the last factor is generated twice and saved as `factor_portfolio`. I will take care of that at some other time.
```{r}
# Define parameters
mode_list <- c("1/n", "relative_signal")
factor_list <- colnames(fx_factors)[-c(1:4)]  # Extract factor names excluding the first four columns
strategy <- "long_short"  # Options: "long_only", "short_only", "long_short"
liquidity_threshold <- 0.008  # Threshold for relative bid-ask spread

# Loop through factor columns
for (i in seq_along(factor_list)) {
  factor_name <- factor_list[i]  # Current factor name
  
  # Construct the factor portfolio
  factor_portfolio <- portfolio_construction(
    data = fx_factors %>%
      select(Currency, date, log_return, Liquidity, all_of(factor_name)),  # Select relevant columns
    n = 10,
    mode = "1/n",
    strategy = strategy,
    liquidity_threshold = liquidity_threshold,
    factor_name = factor_name
  )
  
  
  
  # Dynamically assign the factor portfolio to a variable
  assign(paste0("factor_portfolio.", factor_name), factor_portfolio)
}

# Clean up temporary variables
if (!exists("DEBUG") || !DEBUG) {
  rm(i, factor_name, factor_portfolio)
}

```

Plot individual factors.
```{r}
# Iterate through each factor in factor_list and plot its cumulative log return
for (factor in factor_list) {
  factor_portfolio <- get(paste0("factor_portfolio.", factor))
  plot <- plot_factor_portfolio(factor_portfolio, factor_name = factor)
  print(plot)
}
rm(factor, factor_portfolio, plot)
```

Gives a view over liquditiy.
```{r}
library(ggplot2)

ggplot(fx_factors, aes(x = Currency, y = Liquidity)) +
  geom_boxplot() +
  labs(title = "Box Plots of Liquidity Grouped by Currency",
       x = "Currency",
       y = "Liquidity") +
  theme_minimal(base_size = 10) +  # Increase base font size for the entire plot
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1),  # Rotate x-axis labels
    plot.title = element_text(size = rel(2)),           # Scale title size
    axis.title = element_text(size = rel(1.5)),         # Scale axis title size
    axis.text = element_text(size = rel(1.2))           # Scale axis text size
  )
```

Now that we have the individual factor portfolios, we can move on to computing the multi factor ones. We need weighting schemes for that.
```{r}
factor_weighting <- function(data.timestamp, method = "1/n") {
  if(method == "1/n"){
    weights <- 1/length(unique(data.timestamp$factor_name))
  }
  return(weights)
}

portfolio_merging <- function(factor_combination = c(), method = "1/n") {
  # Initialize multi-factor portfolio with the first factor portfolio
  merged_factor_portfolio <- get(paste0("factor_portfolio.", factor_combination[1]))
  
  # Combine the factor portfolios dynamically
  for (i in 2:length(factor_combination)) {
    # Merge additional factor portfolios
    merged_factor_portfolio <- rbind(merged_factor_portfolio, 
                                     get(paste0("factor_portfolio.", factor_combination[i])))
  }
  
  # Dynamically assign factor weights based on factor_combination
  factor_weighting_storage <- merged_factor_portfolio %>%
    group_by(date) %>%
    reframe(
      # Dynamically match factor_name with factor_combination for weights
      factor_name = factor_combination,
      factor_weight = factor_weighting(data.timestamp = cur_data())
    )
  
  # Replace columns with adjusted values
  merged_factor_portfolio <- merged_factor_portfolio %>%
    left_join(factor_weighting_storage, by = c("date", "factor_name")) %>%
    mutate(
      log_return = log_return * factor_weight,
      weight = weight * factor_weight
    ) %>%
    select(-c(factor_weight))
  
  
  return(merged_factor_portfolio)
}

```

We can now try it out.
```{r}
# Generate factor combinations
factor_combinations_list <- generate_combinations(factor_list, min_constituents = 2, max_constituents = 2)

# Initialize lists to store portfolios and results
portfolios_list <- list()
plots_list <- list()
results_list <- list()

# Loop through each factor combination
for (i in 1:length(factor_combinations_list)) {
  factor_combination <- factor_combinations_list[[i]]
  
  # Construct portfolio and store it in the portfolios list
  portfolios_list[[i]] <- portfolio_merging(factor_combination = factor_combination)
  portfolio_name <- paste(factor_combination, collapse = "_")
  
  # Plot cumulative log return and store the plot
  plots_list[[i]] <- plot_factor_portfolio(
    factor_portfolio = portfolios_list[[i]],
    factor_name = portfolio_name,
    y_min = -0.5,
    y_max = 2.5
  )
  
  # Compute risk metrics and store the result
  risk_metrics <- risk_measure_summary(
    factor_portfolio = portfolios_list[[i]],
    risk_free_rate = risk_free_rate,
    end_date = "2019-12-31"
  )
  
  # Add the factor name to the result
  risk_metrics$factor_name <- portfolio_name
  results_list[[i]] <- risk_metrics
}

# Combine all results into a single data frame
results_df <- do.call(rbind, results_list)

# Print results
print(results_df)

# Optionally print plots
for (plot in plots_list) {
  print(plot)
}

save(portfolios_list, file = "portfolios_list.RData")
save(results_df, file = "results_df.RData")

# Clean up temporary variables
if (!exists("DEBUG") || !DEBUG) {
  rm(plot, factor_combinations_list, results_list, risk_metrics,
     factor_combination, factor_list, i, portfolio_name)
}
```