---
title: "Carry_Strategy"
author: "Maximilian Michlits"
header-includes:
   - \usepackage{bbm}
date: "r Sys.Date()"
output: html_document
---

To Do's:
Set up proper data loading
Inlcude 3 factors that Simion sent (in icloud -> Studium -> Simion, Giorgia "Questions")
Statistical interactions


Starting off by loading libraries and the data.
```{r, include = FALSE}
#load libraries
library(dplyr) # general use
library(forecast) # for autoregression models

#Debug T = on/ F = off
DEBUG <- FALSE
```

We want to calculate the carry of each currency. The Carry $C$, of currency $i$ against USD, at time $t$ and horizon $h\in (0,12)$ in months (the time span into the future for the forward) can be computed as follows: $C_{i_t}^h=Forward_{i_t}^h-Spot_{i_t}^h$

Luckily this is the same as forward spreads which is the standard way of downloading forward data via Bloomberg. Therefore, no further calculations are necessary to find carry.

I use monthly data for the following to reduce noise and speed up computation time. This methodology is further validated by application constraints, as reallocating more often results in extremly high trading costs.

Other factors we load are:
liquidity = (bid-ask) / ((bid+ask)/2), capturing relative spread relative to price.
VIX, which is the implied volatility of the S&P500 for one year, basically markets anticipation of risk for the next year.
```{r}
#load data sets.
load("./Data/fx_log_returns.RData")
load("./Data/fx_carry.RData")
load("./Data/fx_liquidity.RData")
load("./Data/vix.RData")

#Merge Datasets
fx_factors <- left_join(x = fx_log_returns, y = fx_carry, by = join_by(currency, date))
fx_factors <- left_join(x = fx_factors, y = fx_liquidity, by = join_by(currency, date))
fx_factors <- left_join(x = fx_factors, y = vix, by = "date")

#display df
# Display the first and last 50 rows of fx_log_returns
head(fx_factors, 50) %>% 
  knitr::kable()
tail(fx_factors, 50) %>% 
  knitr::kable()

# Clean up!
if (DEBUG != TRUE) {
  rm(fx_carry, fx_liquidity, fx_log_returns, vix)
}
```

#Building models
We try out different models and compare them via statistical tests.

##The interactions model
We use linear models to try predicting log_returns in currency markets.

This function helps us to save diagnostic plots in folders for a model.
```{r}
save_model_diagnostics <- function(model, main_dir = "model_diagnostics") {
  # Automatically set model_name to the name of the variable passed to `model`
  model_name <- deparse(substitute(model))
  
  # Create the main directory if it doesn't exist
  if (!dir.exists(main_dir)) {
    dir.create(main_dir)
  }
  
  # Create a subdirectory for the specific model
  model_dir <- file.path(main_dir, model_name)
  if (!dir.exists(model_dir)) {
    dir.create(model_dir)
  }
  
  file_name_pattern <- file.path(model_dir, paste0(model_name, "_diagnostics"))
  
  # Open a new PNG device for plotting and set PNG device up
  png(paste0(file_name_pattern, ".png"), width = 2000, height = 2000, res = 150)
  par(mfrow = c(3, 2))  # Set layout for 6 plots: 4 diagnostics, ACF, coefficients
  
  # 4 base diagnostic plots, ACF
  plot(model, main = paste("Diagnostic plots for", model_name))
  acf(residuals(model), main = paste("ACF of Residuals for", model_name))

  # Close the PNG device and reset layout
  dev.off()
  par(mfrow = c(1, 1))
  print(sprintf("Saved diagnostic plots for %s to folder: %s", model_name, model_dir))
}

```

Now we run the interactions model.
```{r}
#Setting Parameters
main_dir <- "model_diagnostics"

#Run models on different data sets
#All currencies
LR_int_all <- lm(formula = log_return ~ (carry + liquidity + vix_level)*currency, data = fx_factors)
save_model_diagnostics(model = LR_int_all,
                       main_dir = main_dir)
#summary(LR_int_all)

#DM currencies
fx_factors_DM <- fx_factors %>%
  filter(currency %in% c("EUR", "AUD", "JPY", "GBP", "CAD", "CHF", "NOK", "NZD", "SEK"))
LR_int_DM <- lm(formula = log_return ~ (carry + liquidity + vix_level)*currency, data = fx_factors_DM)
save_model_diagnostics(model = LR_int_DM,
                       main_dir = main_dir)
#summary(LR_int_DM)

#DM_S-Tier currencies
fx_factors_DM_S_Tier <- fx_factors %>%
  filter(currency %in% c("EUR", "AUD", "JPY", "GBP", "CAD", "CHF"))
LR_int_DM_S_Tier <- lm(formula = log_return ~ (carry + liquidity + vix_level)*currency, data = fx_factors_DM)
save_model_diagnostics(model = LR_int_DM_S_Tier,
                       main_dir = main_dir)
summary(LR_int_DM_S_Tier)

```

##Backtesting for interactions model
```{r}
intervall_interactions_model <- function(model_formula, data, max_date){
  
  #seperate the data
  training_data <- data %>%
    filter(date < max_date)
  testing_data <- data %>%
    filter(date == max_date)
  
  #generate a model and let it predict the testing data
  model <- lm(as.formula(model_formula) , data = training_data) 
  testing_data$predicted_values <- predict(model, newdata = testing_data)
  
  return(testing_data)
}

single_return_simulation <- function(testing_data, n){
  #number of long/short positions (example: n=2 means 2 long, 2 short) 
  # Get the top n currencies with the largest predicted log returns
  top_n_largest <- testing_data %>%
    ungroup() %>%  # Remove any grouping that might exist
    arrange(desc(predicted_values)) %>%
    slice_head(n = n)
  top_n_smallest <- testing_data %>%
    ungroup() %>%  # Remove any grouping that might exist
    arrange(predicted_values) %>% #by default ascending
    slice_head(n = n)
  
  #IMPLEMENT WEIGHTING SCEMES LATER ON FOR NOW 1/n
  return(sum(top_n_largest$log_return)-sum(top_n_smallest$log_return))
}


#Finally we generate the time series
#Set parameters
n <- 2 #number of long/short positions (example: n=2 means 2 long, 2 short) 
model_formula <- "log_return ~ (carry + liquidity + vix_level)*currency"
data <- fx_factors_DM_S_Tier
simulation_time_series <- data.frame(date = unique(data$date), log_return = NA)

for(date_index in 12:nrow(simulation_time_series) ){
  date <- simulation_time_series$date[date_index]
  simulation_time_series$log_return[date_index] <- single_return_simulation(testing_data = intervall_interactions_model(model_formula = model_formula, data = data, max_date = date), n = n)
}

```

```{r}
model = LR_int_DM_S_Tier
training_and_testing_data <- fx_factors_DM_S_Tier %>%
  filter(as.Date(date) >= as.Date("2010-01-01") & as.Date(date) < as.Date("2020-01-01"))
  model_formula <- "log_return ~ (carry + liquidity + vix_level)*currency"

```
