---
title: "Carry_Strategy"
author: "Maximilian Michlits"
header-includes:
   - \usepackage{bbm}
date: "`r Sys.Date()`"
output: html_document
---


Starting off by loading libraries and the data. Instead of the S&P500, a different normalizer for risk can be used too. Possibly a synthetic one, that just recieves a standard deviation and then does a random normal?
```{r, include = FALSE}
DEBUG <- TRUE
```

```{r, include = FALSE}
#load libraries
library(dplyr)
library(tidyr)
library(ggplot2)
```
We want to calculate the carry of each currency. The Carry $C$, of currency $i$ against USD, at time $t$ and horizon $h\in (0,12)$ in months (the time span into the future for the forward) can be computed as follows: $C_{i_t}^h=Forward_{i_t}^h-Spot_{i_t}^h$

I use monthly data for the following to reduce noise and speed up computation time. This methodology is further validated by application constraints, as daily reallocation positions results in high trading costs.
```{r}
#load fx spot amd forward rates
load("./Data/fx_spots_and_6m_forwards.RData")
```

As forward rates are described at the spread to the current spot rate in my data set the forward variable turns out to be the carry already.
```{r}
colnames(fx_spots_and_6m_forwards)[4] <- "6m_carry"
fx_spots_and_carry <- fx_spots_and_6m_forwards
rm(fx_spots_and_6m_forwards)
```

#Backtesting
We will now experiment with strategies by utilizing the data in various approaches.

##Simple strategy
The simple strategy can be imposed right away. It does not include any timing variants and will only use the currency with the highest and lowest carry signal as suggested by Brunnermeier, Nagel and Peterson (2008). Moreover, the small transaction volume at PMP makes it even less efficient to trade multiple currencies. This strategy will further be used as a benchmark the measure the impact of a timed alternative.

We need to compare the values for each date to find the currency with the highest (lowest) carry. This will be the investment (funding) currency.
```{r}
max_min_carry_per_date <- fx_spots_and_carry %>%
  group_by(date) %>%
  summarize(
    max_6m_carry = max(`6m_carry`, na.rm = TRUE),
    min_6m_carry = min(`6m_carry`, na.rm = TRUE),
    long_currency = currency[which.max(`6m_carry`)],
    short_currency = currency[which.min(`6m_carry`)]
  )
```

We are plotting the Long (Short) Currencies.
```{r}
# Convert long_currency and short_currency to factors
max_min_carry_per_date$long_currency <- factor(max_min_carry_per_date$long_currency)
max_min_carry_per_date$short_currency <- factor(max_min_carry_per_date$short_currency)

# Plot the currencies with the maximum carry over time
ggplot(max_min_carry_per_date) +
  geom_point(aes(x = date, y = long_currency, color = "Long Currency"), size = 2) +
  geom_point(aes(x = date, y = short_currency, color = "Short Currency"), size = 2) +
  labs(title = "Currencies with Maximum and Minimum Carry Over Time",
       x = "Date",
       y = "Currency",
       color = "Legend") +
  scale_color_manual(values = c("Long Currency" = "red", "Short Currency" = "blue")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

We can now use return data to compute returns of the simple strategy.
```{r}
#Load the data frame
load("./Data/fx_log_returns.RData")

#Take max_min_carry_per_date and merge with respective log returns.
max_min_carry_per_date <- max_min_carry_per_date[c("date","long_currency","short_currency")]

# Ensure the currency columns in fx_log_returns are factors to match max_min_carry_per_date
fx_log_returns <- fx_log_returns %>%
  mutate(currency = as.factor(currency))

# Join the data frames
simple_strategy_df <- max_min_carry_per_date %>%
  left_join(fx_log_returns, by = c("date" = "date", "long_currency" = "currency")) %>%
  rename(long_log_return = log_return) %>%
  left_join(fx_log_returns, by = c("date" = "date", "short_currency" = "currency")) %>%
  rename(short_log_return = log_return)

# Display the structure of the joined data
str(simple_strategy_df)

if(DEBUG != TRUE){
  rm(max_min_carry_per_date)
}
```
Compute monthly strategy returns.
```{r}
#Properly account for short exposure
simple_strategy_df$short_log_return <- simple_strategy_df$short_log_return*(-1)

#Calculate monthly total returns and cumulative total returns. We deal with NAs by "not investing" resulting in 0 total returns in the given month.
simple_strategy_df <- simple_strategy_df %>%
  mutate(total_log_return = long_log_return + short_log_return) %>%
  mutate(total_log_return = replace_na(total_log_return, 0)) %>%
  mutate(cumulative_log_return = cumsum(total_log_return))
```

We plot the simple strategy, because... why not!
```{r}
ggplot(simple_strategy_df, aes(x = date, y = cumulative_log_return)) +
  geom_line() +
  labs(title = "Cumulative Log Returns Over Time",
       x = "Date",
       y = "Cumulative Log Return") +
  theme_minimal()
```

```{r}

```