---
title: "Carry_Strategy"
author: "Maximilian Michlits"
header-includes:
   - \usepackage{bbm}
date: "r Sys.Date()"
output: html_document
---

To Do's:
Set up proper data loading
Inlcude 3 factors that Simion sent (in icloud -> Studium -> Simion, Giorgia "Questions")



Starting off by loading libraries and the data.
```{r, include = FALSE}
#load libraries
library(dplyr) # general use
library(forecast) # for autoregression models

#Debug T = on/ F = off
DEBUG <- FALSE
```

We want to calculate the carry of each currency. The Carry $C$, of currency $i$ against USD, at time $t$ and horizon $h\in (0,12)$ in months (the time span into the future for the forward) can be computed as follows: $C_{i_t}^h=Forward_{i_t}^h-Spot_{i_t}^h$

Luckily this is the same as forward spreads which is the standard way of downloading forward data via Bloomberg. Therefore, no further calculations are necessary to find carry.

I use monthly data for the following to reduce noise and speed up computation time. This methodology is further validated by application constraints, as reallocating more often results in extremly high trading costs.

Other factors we load are:
liquidity = (bid-ask) / ((bid+ask)/2), capturing relative spread relative to price.
VIX, which is the implied volatility of the S&P500 for one year, basically markets anticipation of risk for the next year.
```{r}
#load data sets.
load("./Data/fx_log_returns.RData")
load("./Data/fx_carry.RData")
load("./Data/fx_liquidity.RData")
load("./Data/vix.RData")

#Merge Datasets
fx_factors <- left_join(x = fx_log_returns, y = fx_carry, by = join_by(currency, date))
fx_factors <- left_join(x = fx_factors, y = fx_liquidity, by = join_by(currency, date))
fx_factors <- left_join(x = fx_factors, y = vix, by = "date")

#display df
# Display the first and last 50 rows of fx_log_returns
head(fx_factors, 50) %>% 
  knitr::kable()
tail(fx_factors, 50) %>% 
  knitr::kable()

# Clean up!
if (DEBUG != TRUE) {
  rm(fx_carry, fx_liquidity, fx_log_returns, vix)
}
```

#Buildering linear models
We use linear models to try predicting log_returns in currency markets.

##Defining functions for generating and analyzing linear models
We use this function to generate all model combinations for a list of dependend and independent variables.
```{r}
# Function to generate all combinations of independent variables and dependent variables
generate_model_combinations <- function(dependent_var, independent_vars, subset_models = NA) {
  # Initialize an empty list to store model formulas
  model_combinations <- list()
  
  # If subset_models is not NA, apply the provided conditions
  if (!is.na(subset_models[1])) {
    for (subset in subset_models) {
      # Split the subset string to get the independent variables and condition
      subset_parts <- unlist(strsplit(subset, "_"))
      
      # The last part is the condition (e.g., "JPY"), and the rest are the independent variables
      condition <- subset_parts[length(subset_parts)]
      
      # Generate all combinations of independent variables for each subset
      for (i in 1:length(independent_vars)) {
        combos <- combn(independent_vars, i, simplify = FALSE)
        
        # For each combination, create the model name and formula with the subset condition
        for (combo in combos) {
          # Create a model name (e.g., "lm_log_return_carry_liquidity_JPY")
          model_name <- paste("lm", dependent_var, paste(combo, collapse = "_"), condition, sep = "_")
          
          # Create a formula string that applies the condition to both dependent and independent variables
          dependent_string <- paste(dependent_var, "[currency == '", condition, "']", sep = "")
          independent_string <- paste(lapply(combo, function(var) paste(var, "[currency == '", condition, "']", sep = "")), collapse = " + ")
          
          # Combine into a formula string
          formula_string <- paste(dependent_string, "~", independent_string)
          
          # Add the formula string to the list
          model_combinations[[model_name]] <- formula_string
        }
      }
    }
  } else {
    # If subset_models is NA, generate all combinations of independent variables without a condition
    for (i in 1:length(independent_vars)) {
      # Get all combinations of i independent variables
      combos <- combn(independent_vars, i, simplify = FALSE)
      
      # For each combination, create the formula and model name
      for (combo in combos) {
        # Create a model name (e.g., "lm_dependent_carry_liquidity")
        model_name <- paste("lm", dependent_var, paste(combo, collapse = "_"), sep = "_")
        
        # Create a formula string (e.g., "dependent_var ~ carry + liquidity")
        formula_string <- paste(dependent_var, "~", paste(combo, collapse = " + "))
        
        # Add the formula string to the list
        model_combinations[[model_name]] <- formula_string
      }
    }
  }
  
  return(model_combinations)
}
```

To run all the models in a list (the returned object of generate_model_combinations) we define a function that runs all models in a list of model-formulas.
```{r}
fit_models <- function(models = list(), data = NA) {
  # Initialize an empty list to store the fitted models
  model_list <- list()
  
  # Loop through each model in model_combinations
  for (model_name in names(model_combinations)) {
    # Get the formula as a string
    formula_string <- model_combinations[[model_name]]
    
    # Fit the linear model using the formula and the provided dataset
    model <- lm(as.formula(formula_string), data = data)
    
    # Store the fitted model in the list
    model_list[[model_name]] <- model
  }
  
  # Return the list of fitted models
  return(model_list)
}
```

And a function that saves diagnostic plots in folders for a list of models.
```{r}
save_model_diagnostics <- function(model_list, main_dir = "model_diagnostics", subset_names = NA) {
  # Create the main directory if it doesn't exist
  if (!dir.exists(main_dir)) {
    dir.create(main_dir)
  }
  
  # Check if we are using currency-specific models or pooled models
  if (!is.na(subset_names[1])) {
    # Loop over each subset in the subset_names list (e.g., "JPY", "USD")
    for (subset in subset_names) {
      # Create a subdirectory for each subset (e.g., "JPY_models", "USD_models")
      sub_dir <- paste0(subset, "_models")
      output_dir <- file.path(main_dir, sub_dir)
      
      # Create the subdirectory if it doesn't exist
      if (!dir.exists(output_dir)) {
        dir.create(output_dir, recursive = TRUE)
      }
      
      # Now generate diagnostic plots only for models that match the current subset (currency)
      for (model_name in names(model_list)) {
        # Check if the model name ends with the current subset (e.g., "JPY", "USD")
        if (grepl(paste0("_", subset, "$"), model_name)) {
          # Extract the linear model
          model <- model_list[[model_name]]
          
          # Define the file name for the diagnostic plots
          file_name <- file.path(output_dir, paste0(model_name, "_diagnostics.png"))
          
          # Open a new PNG device
          png(file_name, width = 1200, height = 1200)
          
          # Set a proper plot layout
          par(mfrow = c(3, 2))
          
          # Plot the model diagnostics
          plot(model, main = paste("Diagnostic plots for", model_name))
          
          # Plot the autocorrelation function of residuals
          acf(residuals(model), main = paste("ACF of Residuals for", model_name))
          
          # Close the PNG device
          dev.off()
          
          print(paste("Saved diagnostic plots for", model_name, "to", file_name))
        }
      }
    }
  } else {
    # Case when subset_names is NA (pooled data)
    # Create a pooled directory
    pooled_dir <- file.path(main_dir, "pooled_models")
    
    if (!dir.exists(pooled_dir)) {
      dir.create(pooled_dir)
    }
    
    # Generate diagnostic plots for pooled models
    for (model_name in names(model_list)) {
      # Extract the linear model
      model <- model_list[[model_name]]
      
      # Define the file name for the diagnostic plots
      file_name <- file.path(pooled_dir, paste0(model_name, "_diagnostics.png"))
      
      # Open a new PNG device
      png(file_name, width = 1200, height = 1200)
      
      # Set a proper plot layout
      par(mfrow = c(3, 2))
      
      # Plot the model diagnostics
      plot(model, main = paste("Diagnostic plots for", model_name))
      
      # Plot the autocorrelation function of residuals
      acf(residuals(model), main = paste("ACF of Residuals for", model_name))
      
      # Close the PNG device
      dev.off()
      
      print(paste("Saved diagnostic plots for", model_name, "to", file_name))
    }
  }
  
  # Reset the plotting layout back to default (1x1)
  par(mfrow = c(1, 1))
}

```

##Setting Parameters
```{r}
dependent_var <- c("log_return")
independent_vars <- c("carry", "liquidity", "vix_level")
```

###Pooled currency models
We can now run the pooled currency models and save them
```{r}
#We get our combinations
model_combinations <- generate_model_combinations(dependent_var, independent_vars)

#Fit the models
model_list_pooled <- fit_models(models = model_combinations,
                                data = fx_factors)

#And save the plots
main_dir <- "model_diagnostics"
save_model_diagnostics(model_list = model_list_pooled,
                       main_dir = main_dir)
```

###Currency specific models
As we can see, due to the large number of currencies and some individual properties we cannot assume ???. To resolve this we try running currency specific models and analyze the residuals.
```{r}
#We get our combinations
model_combinations <- generate_model_combinations(dependent_var,
                                                  independent_vars,
                                                  subset_models = unique(fx_factors$currency))
#Fit the models
model_list_currency_specific <- fit_models(models = model_combinations, 
                                           data = fx_factors)

#And save the plots
save_model_diagnostics(model_list = model_list_currency_specific,
                       main_dir = main_dir,
                       subset_names = unique(fx_factors$currency))
```

```{r}

```

```{r}

```
